---
title: "Mountain Yellow-legged Frog Population in the Sierra Lakes Inventory Project"
author: "Teague Tran"
date: "1/23/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = TRUE, message = FALSE, warning = FALSE}
#Load necessary packages
library(tidyverse)
library(janitor)
library(palmerpenguins)
library(here)

#For ggplot customization
library(readxl)
library(gghighlight)
library(lubridate)
library(plotly)

#to make compound figures 
library(patchwork)

#read in .xlsx file
sierra_amphibians <- read_excel(here("data", "sierra_amphibians.xlsx"))
```

```{r, echo = TRUE, message = FALSE, warning = FALSE}
#graph total mountain yellow-legged frog count each year by life stage across all bodies
# create "year" column 
sierra_amphibians <- sierra_amphibians %>% 
  mutate(date = ymd(survey_date)) %>% 
  mutate(year = year(date))
# select the required data (year and life stage)
ylfrog_count <- sierra_amphibians %>% 
  select(survey_date, amphibian_species, amphibian_life_stage, year) %>% 
  filter(amphibian_species == "RAMU", amphibian_life_stage != "EggMass") %>% 
  group_by(year, amphibian_life_stage) %>% 
  count()

ylfrog_count_graph <- 
  ggplot(data = ylfrog_count, aes(x = year, y = n, color = amphibian_life_stage)) +
  geom_line() + 
  labs(x = "Year", y = "Amphibian Count", title = "Yellow-legged Frog Count of Different Life Stages (1995-2002)", color = "Life Stages")

```
```{r, echo = TRUE, message = FALSE, warning = FALSE}
# create a column graph containing total counts of combined adult and subadult yl frogs observed in the 5 lakes with the greated counts
ylfrog_top5 <- sierra_amphibians %>% 
  select(lake_id, amphibian_species, amphibian_life_stage) %>% 
  filter(amphibian_species == "RAMU", amphibian_life_stage == "Adult"| amphibian_life_stage == "SubAdult") %>% 
  count(lake_id) %>% 
  slice_max(n, n=5)

#make lake_id into character and add "Lake" to the beginning of each one
ylfrog_top5 <- ylfrog_top5 %>% 
  mutate(Lakeid = as.character(lake_id)) %>% 
  mutate(Lake = "Lake") %>% 
  unite("LakeID", Lake:Lakeid, sep = " ") %>% 
  mutate(LakeID = fct_reorder(LakeID, n))

#create a column graph
yl_frog_top5_graph <-
  ggplot(data = ylfrog_top5, aes(x = LakeID, y = n)) +
  geom_col() +
  labs(x = " ", y = "Amphibian Count", title = "Top 5 lakes with the highest observed counts of mountain yellow-legged frog")
```

```{r, echo = TRUE, message = FALSE, warning = FALSE}
#Combine the two graphs I just made into a single compound figure using {patchwork} package

combined_graph <- ylfrog_count_graph / yl_frog_top5_graph

combined_graph

```

Figure 1. Mountain yellow-legged frog (*Rona muscosa*) count data from the Sierra Lakes Inventory Project (1995-2002) are shown. The top graph displays total mountain yellow-legged frog count throughout the entire study period and area but excludes frogs in the egg mass life stage. The bottom graph displays the counts of the top 5 lakes where adult and subadult mountain yellow-legged frog counts are the highest. Source: [Knapp et al. 2020](https://portal.edirepository.org/nis/metadataviewer?packageid=edi.577.2)


