---
title: "a1_task2_Teague_Tran"
author: "Teague Tran"
date: "1/23/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,echo = FALSE, message = FALSE, warning = FALSE}
#Set up section
library(tidyverse)
library(janitor)
library(palmerpenguins)
library(here)

#for PCA
library(ggfortify)

#For ggplot customization
library(readxl)
library(gghighlight)
library(lubridate)
library(plotly)

#to make compound figures 
library(patchwork)

#read in .csv file
world_env_vars <- read_csv(here("data", "world_env_vars.csv"))
```

### Introduction: 

This short study is to determine if there are relationships between the world's tree canopy cover, mean annual temperature, cloudinessm, and cropland cover. The data used in this study provided by [@zander_venter](https://www.kaggle.com/zanderventer/datasets) on Kaggle, who acquired the data through [Google Earth Engine](https://earthengine.google.com/). To determine the relationships between the variables, we performed a principal components analysis, which captured approximately 77% of the variance when comparing these variables. 


```{r,echo = TRUE, message = FALSE, warning = FALSE}
#create new columns with better variable names to make the graph more professional 
world_env_vars <- world_env_vars %>% 
  mutate("Cropland Cover" = cropland_cover, 
         "Tree Canopy Cover" = tree_canopy_cover,
         "Mean Annual Temperature" = temp_mean_annual,
         "Cloudiness" = cloudiness)
#select variables that I want to include in my PCA biplot
env_vars_PCA <- world_env_vars %>%
  select("Cropland Cover", "Tree Canopy Cover", "Mean Annual Temperature", "Cloudiness") %>% 
  drop_na() %>% 
  scale() %>% #allows us to compare variables evenly
  prcomp() #run PCA on data
```

```{r,echo = TRUE, message = FALSE, warning = FALSE}
#create biplot
env_vars_complete <- world_env_vars %>% 
  drop_na(cropland_cover, tree_canopy_cover, temp_mean_annual, cloudiness)

autoplot(env_vars_PCA,
         data = env_vars_complete, 
         colour = 'black',
         loadings = TRUE,
         loadings.label = TRUE) + 
  theme_minimal() +
  labs(title = "Relationship between different envrionmental variables in the world")
```

Figure 1. Approximately 77% (PC1 = 42.91% and PC2 = 34.82%) of the variance in the principal components analysis is capture in this biplot. Observations appear to be evenly spread, with a slight cluster between the Mean Annual Temperature arrow and the Tree Canopy Cover arrow.  

### Summary:

- The principal components in this biplot capture approximately 77% of variance found in the data (PC1 = 42.91% and PC2 = 34.82%)
- Mean annual temperature has a strong negative correlation with cropland cover
- Mean temperature and cropland cover have a weak correlation with tree canopy cover and cloudiness
- Tree canopy cover and cloudiness are positively correlated with each other


